---
author: "Bella Garcia"
title: "EPI 590R R Project"
format: html
---

***For my final project I am using the COVID-19 testing cohort dataset from Amrom E. Obstfeld that was collected from the Children's Hospital of Pennsylvania. This dataset was found in the {medicaldata} package in R that was developed by Peter Higgins.***

```{r}
library(readr)
library(tidyverse)
library(gtsummary)
library(dplyr)
library(broom.helpers)
library(cards)
library(cardx)
library(broom)
library(medicaldata)
```

```{r}
#| message: false
data(package="medicaldata")
covid <- medicaldata::covid_testing
class(covid)
covid <- as.data.frame(covid)

write_csv(covid, here::here("Data", "raw", "covid.csv"))


#cleaning data
covid_clean <- covid |>
  mutate(clinic_name = case_when(str_detect(clinic_name, "onc") ~ "oncology",str_detect(clinic_name, "mri") ~ "MRI", str_detect(clinic_name, "onc") ~ "oncology", str_detect(clinic_name, "rad") ~ "diagnosis radiology",str_detect(clinic_name, "urg") ~ "urgent care",
           str_detect(clinic_name, "lab") ~ "laboratory",str_detect(clinic_name, "ntwk") ~ "care network",
           str_detect(clinic_name, "hosp") ~ "hospital",
           str_detect(clinic_name, "ward") ~ "ward",
           TRUE ~ clinic_name 
         ),
         clinic_name = fct_lump(clinic_name, n = 10))

write_rds(covid_clean, here::here("Data", "clean", "covid_clean.rds"))
write_csv(covid_clean, here::here("Data", "clean", "covid_clean.csv"))
```

```{r}
#| label: tbl-first
#| tbl-cap: "Descriptive Statistics"
  table1 <- tbl_summary(
  covid_clean,
  by = gender,
  include = c(age, result, pan_day, test_id, clinic_name, payor_group, demo_group, patient_class),
  label = list(
    payor_group ~ "Paying method",
    pan_day ~ "# of days tested after pandemic start",
    test_id ~ "Type of test",
    clinic_name ~ "Clinic",
    demo_group ~ "Subject groups", 
    patient_class ~ "Subject disposition" 
  ),
  missing_text = "Missing")   |> 
 	add_overall(col_label = "**Total**") |> 
  bold_labels() 
 table1 
mean_age <- inline_text(table1,variable = "age", column="stat_0")
```

See @tbl-first for the descriptive statistics of the data

The IQR age of those taking a COVID-19 test is `r mean_age`

```{r}
#| label: tbl-lm
#| tbl-cap: "This is a regression model"
regressionmodel <- tbl_uvregression(
  covid_clean,
  y=age, 
  method = lm,
  include = c(age, clinic_name,result, gender, payor_group, demo_group, patient_class))
regressionmodel

```

```{r}
#create function of sample std
stdfunction <- function(col_rec_tat) {
  n <- length(col_rec_tat)            
  mean <- sum(col_rec_tat) / n          
  diffsq <- (col_rec_tat - mean)^2       
  variance <- sum(diffsq) / (n - 1)      
  std_val <- sqrt(variance)                      
  
  return(std_val)
}

std <-sd(covid_clean$col_rec_tat, na.rm=TRUE)


stdfunction(covid_clean$col_rec_tat)



```

See @tbl-lm for the regression model for this data

The standard deviation of the time elapsed between collect time and receive time is (`r std`)

```{r}
#| label: fig-hist
#| fig-cap: "This is a histogram"
figure1 <- hist(covid_clean$age,main="Histogram of Age", 
     xlab = "Age", 
     ylab = "Frequency", 
     col = "coral",
     border = "black")

```

See @fig-hist for a histogram of the age in the COVID dataset
